Revisiting the hydrodynamic modulation of short surface waves by longer waves

### 2024-10-15 paper.tex
Near-final edits

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that affects remote sensing, the
interpretation of \textit{in situ} wave measurements, and numerical wave
forecasting.
In this paper, we revisit the linear (Airy) wave theory and derive higher-order
steady solutions for the change of short-wave wavenumber, action density, and
gravitational acceleration due to the presence of longer waves.
We validate the analytical solutions with numerical simulations of the full wave
crest and action conservation equations.
The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
steepness modulation significantly deviate from the linear analytical solutions
of \citet{longuet1960changes}, and are similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
We further examine the result of \citet{peureux2021unsteady} who found through
numerical simulations that the short-wave modulation grows unsteadily with
each long-wave passage.
We show that this unsteady growth only occurs only for homogeneous initial
conditions as a special case and does not generally occur, for example in more
realistic long-wave groups.
The combined analytical and numerical framework presented here allows
quantifying the importance of various terms in the full wave conservation
equations.
The proposed steady solutions are a good approximation of the fully
nonlinear numerical solutions in long-wave steepness up to $\approx 0.2$.
As a corollary, except for a subset of initial conditions, the solutions to
the fully nonlinear crest-action conservation equations are mostly steady in
the reference frame of the long waves.
\end{abstract}

@@

\begin{equation}
\label{eq:phi_L}
\phi_L = \dfrac{a_L \omega_L}{k_L} e^{k_L z} \sin{\psi} + \mathcal{O}(\varepsilon_L^4)
\end{equation}
where $z$ is the vertical distance from the mean water level that is positive
upwards.
(\ref{eq:phi_L}) is accurate to the third order in $\varepsilon_L$.
The long-wave induced horizontal and vertical orbital velocities, evaluated at
the wave surface $z = \eta_L$, are:

\begin{equation}
\label{eq:U_L}
U = a_L \omega_L e^{k_L z} \cos{\psi} + \mathcal{O}(\varepsilon_L^4)
\end{equation}

\begin{equation}
\label{eq:w_L}
W = a_L \omega_L e^{k_L z} \sin{\psi} + \mathcal{O}(\varepsilon_L^4)
\end{equation}
Evaluating these velocities at the wave surface $z = \eta_L$ rather than at
the mean water level $z=0$ is a departure from the linear theory, making the
treatment of the long waves here as weakly nonlinear.
As \citet{zhang1990evolution} point out, this is appropriate to do when the
amplitude of the long waves is of the similar magnitude or large than the
wavelength of the short waves.

@@

\begin{equation}
\label{eq:k_short_lhs}
\frac{\widetilde{k}}{k} = 1 + \varepsilon_L \cos{\psi}
\end{equation}
Coming from the conservation of crests, as opposed to the velocity potential
superposition, the solution by \citet{longuet1960changes} requires two
approximations:
First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
second, further truncating the Taylor expansion series beyond the first-order
term.
These approximations underestimate the short-wave modulation magnitude
(Fig. \ref{fig:analytical_solutions_ak0.1}a).
At the wave crest, (\ref{eq:k_short_exact}) yields the wavenumber modulation
that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of
\citet{longuet1960changes} for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
Although not immediately obvious from the expressions, the modulated wavenumber
(\ref{eq:k_short_lhs}) is conserved across the long-wave phase, whereas
the higher-order solution (\ref{eq:k_short_exact}) is not.
This is because combining the orbital velocities evaluated at $z=\eta$ and the
linearised conservation of wave crests in (\ref{eq:wavenumber}) results in
excess short-wave wavenumber being created.
The solution of (\ref{eq:wavenumber}) must be conservative across the long-wave
phase, as we will see later from the numerical simulations.

@@

\begin{equation}
\label{eq:gravity_modulation_linear_no_curvature}
\frac{\widetilde{g}}{g} = \left( 1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}} \right)
\end{equation}
Further, if we evaluate the orbital accelerations at the mean water level $z=0$,
this further simplifies to:

\begin{equation}
\label{eq:gravity_modulation_linear_no_curvature_mean_level}
\frac{\widetilde{g}}{g} = \left( 1 - \varepsilon_L \cos{\psi} \right)
\end{equation}
which is the form used by \citet{peureux2021unsteady}, for example.

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_effective_gravities.png}
\caption{
  (a) Analytical solutions for the effective gravitational acceleration
  modulation by long waves, as function of long-wave phase, for $\varepsilon_L = 0.2$.
  Black is for linear wave evaluated at $z=0$,
  blue and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and without curvature effects,
  and orange and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and with curvature effects.
  (b) As in (a) but normalised by the modulation of the linear wave evaluated at $z=0$, $g_1$.
}
\label{fig:effective_gravities}
\end{figure}

It is instructive to also examine the mean, maximum, and minimum values of the
effective gravity as function of long-wave steepness $\varepsilon_L$
(Fig. \ref{fig:effective_gravities_mean_max_min}).
The long-wave averaged effective gravity of short waves remains equal to the
resting value of $g$ for all values of $\varepsilon_L$, except if the effects
of curvature are considered, in which case the average effective gravity is
reduced.
This reduction, in case of a long Stokes wave, is $\approx 3\%$ for
$\varepsilon_L = 0.2$ and $\approx 14\%$ for $\varepsilon_L = 0.4$.
As the effective gravity minima and maxima occur at crests and troughs,
respectively, the curvature affects the phase-averaged values but not the extremes.
The effective gravity reduction at the crests is $\approx 11\%$, $25\%$, and $66\%$
for $\varepsilon_L = 0.1$, $0.2$, and $0.4$, respectively.
These are similar to the fully-nonlinear numerical estimates of
\citet{longuet1986eulerian,longuet1987propagation}.
Note that these are strictly Eulerian values, evaluated at fixed positions in
space.
As \citet{longuet1986eulerian} pointed out, the effective gravity in the
Lagrangian frame (\textit{i.e.}, following the surface particles) is smoother
and less amplified because the moving particles spend a longer time riding near
the crests and less time returning through the troughs.
However, when we solve the short-wave crest and action conservation equations
in an Eulerian frame, it is the Eulerian effective gravity that must be applied
to conserve the short-wave properties across the long-wave phase.

@@

\begin{equation}
\label{eq:wave_amplitude_modulation}
\dfrac{\widetilde{a}}{a} = \sqrt{
  \dfrac{g}{\widetilde{g}}
  \dfrac{\widetilde{\sigma}}{\sigma}
  \dfrac{\widetilde{N}}{N}}
=
  \left( \dfrac{g}{\widetilde{g}} \right)^{0.25}
  \left( \dfrac{\widetilde{k}}{k} \right)^{0.25}
  \left( \dfrac{\widetilde{N}}{N} \right)^{0.5}
\end{equation}
To the first order in $\varepsilon_L$, and using (\ref{eq:k_short_exact})
and (\ref{eq:wave_action_modulation}), the above simplifies to the result of
\citet{longuet1960changes}:

\begin{equation}
\label{eq:wave_amplitude_modulation_order1}
\dfrac{\widetilde{a}}{a} = 
  \left( \dfrac{g}{\widetilde{g}} \right)^{0.25}
  \left( 1 + \varepsilon_L \right)^{0.75}
  \approx \frac{(1 + \varepsilon_L)^{0.75}}{(1 - \varepsilon_L)^{0.25}}
  \approx 1 + \varepsilon_L
\end{equation}

The amplitude and steepness modulation of the short waves as a function of the
long-wave phase for $\varepsilon_L = 0.1$, $0.2$, and $0.4$ are shown and compared
to the \citet{longuet1960changes} solutions in Figs.
\ref{fig:analytical_solutions_ak0.1}, \ref{fig:analytical_solutions_ak0.2},
and \ref{fig:analytical_solutions_ak0.4}, respectively.
Also shown are the modulation of the short-wave intrinsic frequency and phase
speed in panels (e) and (f), respectively.
As the wavenumber and gravity modulations are out of phase, with the latter
being somewhat larger in magnitude, the overall intrinsic frequency of the
short waves is reduced across the entire long-wave phase, with highest reduction
at the long-wave crests.
For the short-wave phase speed modulation, the wavenumber and gravity modulations
work in tandem and cause the short-wave phase speed to be reduced on the long-wave
crests and increased in the troughs.
We discuss the implications of this effect later in \S\ref{section:discussion}
in the context of wind input to short waves.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.2.png}
\caption{As Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.2$.}
\label{fig:analytical_solutions_ak0.2}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.4.png}
\caption{As Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.4$.}
\label{fig:analytical_solutions_ak0.4}
\end{figure}

@@

Why does this unsteady growth of short-wave action occur?
\citet{peureux2021unsteady} explain that the change in advection velocity due to
the modulation of the short-wave group speed yields an additional amplification
of the short-wave action.
That is, the inhomogeneity term $N \dfrac{\partial C_g}{\partial x}$ is
responsible for the instability.
However, the instability only occurs if the short waves are initialised as a
uniform field.
If they are initialised instead as a periodic function of the long-wave phase
such that their wavenumber (and optionally, action) is higher on the crests
and lower in the troughs, akin to the prior steady solutions, the instability
vanishes and the short-wave modulation pattern locks-in to the long-wave
crests.

@@

Why is it, then, that the solution for short-wave modulation is not only
quantitatively, but also qualitatively, different depending on the short-wave
initial conditions?
It may at first seem that there is something special about the uniform initial
conditions that makes its solutions grow unsteadily and indefinitely.
\citet{peureux2021unsteady} described this scenario as "...the sudden appearance
of a long wave perturbation in the middle of a homogeneous short wave field".
However, it is not obvious whether such a scenario is realistic as waves tend to
disperse into groups, and the time scales of incoming swell fronts are much
longer than those of short-wave generation by wind.
To test whether the sudden onset of long-wave forcing causes the unsteady growth,
we repeat the numerical simulation of Fig.
\ref{fig:modulation_3panel_infinite}, except for a linear ramp applied to $a_L$
(and consequently, the other long-wave properties including the orbital
velocities and the local steepness) for the first five periods:
$a_L(t) = min(a_L, a_L t / \left(5 T_L\right))$, where $T_L$ is the long-wave period.
This result is shown in Fig. \ref{fig:modulation_3panel_ramp}.
The key distinction that arises with the introduction of the ramp is that the
modulation pattern of short waves is now locked-in to the long-wave crests,
and their indefinite steepening vanishes.
The stabilisation effect of the long-wave ramp is more readily assessed in
Fig. \ref{fig:unsteady_growth_timeseries}, which shows the maximum short-wave
steepness modulation as function of time over 30 long-wave periods, for the
infinite long-wave train case and the case of applying a linear ramp.
Comparing the wave action propagation ($C_g \frac{\partial N}{\partial x}$) and
inhomogeneity ($N \frac{\partial C_g}{\partial x}$) tendencies reveals that
after only one long-wave period, the propagation and inhomogeneity tendencies
are unbalanced and amplify the steepening of short waves
(Fig. \ref{fig:inhomogeneity_tendencies}).
\citet{peureux2021unsteady} correctly pointed out that the unsteady growth of
short-wave steepness occurs due to the resonance of the inhomogeneity
tendency with the short-wave action perturbation.
However, in the case of a gentle ramp-up of the long-wave forcing, the
propagation and inhomogeneity tendencies mostly balance each other and allow
the short-wave steepening to lock-in to the long-wave crests.

@@

In this paper we revisited the problem of short-wave modulation by long waves
from a linear wave theory perspective.
Using the linearised wave crest and action conservation equations coupled with
nonlinear solutions for the effective gravity, we derived steady analytical
solutions for the modulation of short wave wavenumber, amplitude, steepness,
intrinsic frequency, and phase speed.
The latter two have not been previously examined in the context of short-wave
modulation, but may be important for correctly calculating the wind input into
short waves, which is proportional to the wind speed relative to the wave
celerity.
The approximate steady solutions yield higher wavenumber and amplitude
modulation than that of \citet{longuet1960changes}, and somewhat lower than 
the numerical solutions of \citet{longuet1987propagation} and
\citet{zhang1990evolution}.
This is mainly due to the analytical solution requiring the crest and action
conservation equations to be linearised, and the nonlinearity arising due
to the evaluation of short waves on the surface of the long wave (as opposed to
the mean water level) and due to nonlinear effective gravity.

The numerical solutions of the full wave crest and action conservation equations
are consistent with the analytical solutions for $\varepsilon_L \lesssim 0.2$.
The simulations also reveal that the unsteady growth of short-wave steepness
reported by \citet{peureux2021unsteady} is due to the sudden appearance of
long waves in a homogeneous short-wave field, a scenario that is unlikely to
occur in the open ocean.
A more likely condition such as that of a long-wave groups, are only mildly
destabilising for the short waves.
These results suggest that for mildly-to-moderately steep short waves,
the analytical steady solutions presented here are sufficient to describe the
modulation of short waves by long waves.
It may be worthwhile to revisit a more systematic inclusion of the hydrodynamic
modulation effects in phase-averaged spectral wave models, beyond the simple
cumulative mean squared slope for wave dissipation, but also for modulating
the dispersion and wind input into short waves.

\backsection[Acknowledgements]{
Thanks to Nathan Laxague, Fabrice Ardhuin, Nick Pizzo, and Peisen Tan for the
helpful discussions that improved this paper.
}

\backsection[Funding]{
  Milan Curcic was partially supported by the National Science Foundation
  under the grant number AGS-1745384.
}

>>>>

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that affects remote sensing, the
interpretation of \textit{in situ} wave measurements, and numerical wave
forecasting.
In this paper, we revisit the linear (Airy) wave theory and derive higher-order
steady solutions for the change of short-wave wavenumber, action density, and
gravitational acceleration due to the presence of longer waves.
We validate the analytical solutions with numerical simulations of the full wave
crest and action conservation equations.
The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
steepness modulation significantly deviate from the linear analytical solutions
of \citet{longuet1960changes}, and are similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
The short-wave steepness modulation is attributed to be 5/8 due to the
wavenumber, 1/4 due to the wave action, and 1/8 due to the effective gravity.
We further examine the result of \citet{peureux2021unsteady} who found through
numerical simulations that the short-wave modulation grows unsteadily with
each long-wave passage.
We show that this unsteady growth only occurs only for homogeneous initial
conditions as a special case and does not generally occur, for example in more
realistic long-wave groups.
The combined analytical and numerical framework presented here allows
quantifying the importance of various terms in the full wave conservation
equations.
The proposed steady solutions are a good approximation of the fully
nonlinear numerical solutions in long-wave steepness up to $\approx 0.2$.
As a corollary, except for a subset of initial conditions, the solutions to
the fully nonlinear crest-action conservation equations are mostly steady in
the reference frame of the long waves.
\end{abstract}

@@

\begin{equation}
\label{eq:phi}
\phi = \dfrac{a_L \omega_L}{k_L} e^{k_L z} \sin{\psi} + \mathcal{O}(\varepsilon_L^4)
\end{equation}
where $z$ is the vertical distance from the mean water level that is positive
upwards.
(\ref{eq:phi}) is accurate to the third order in $\varepsilon_L$ in deep water
because the second- and third-order terms in the Stokes expansion series are
exactly zero.
The long-wave induced horizontal and vertical orbital velocities are:

\begin{equation}
\label{eq:U_L}
U =
\frac{\partial \phi}{\partial x} =
a_L \omega_L e^{k_L z} \cos{\psi} + \mathcal{O}(\varepsilon_L^4)
\end{equation}

\begin{equation}
\label{eq:w_L}
W =
\frac{\partial \phi}{\partial z} =
a_L \omega_L e^{k_L z} \sin{\psi} + \mathcal{O}(\varepsilon_L^4)
\end{equation}
Evaluating these velocities at the wave surface $z = \eta_L$ rather than at
the mean water level $z=0$ is a departure from the linear theory, making the
treatment of the long waves here as weakly nonlinear.
As \citet{zhang1990evolution} point out, this is appropriate to do when the
amplitude of the long waves is of the similar magnitude or large than the
wavelength of the short waves.

@@

\begin{equation}
\label{eq:k_short_lhs}
\frac{\widetilde{k}}{k} = 1 + \varepsilon_L \cos{\psi}
\end{equation}
Coming from the conservation of crests, as opposed to the velocity potential
superposition, the solution by \citet{longuet1960changes} requires two
approximations:
First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
second, further truncating the Taylor expansion series beyond the first-order
term.
These approximations underestimate the short-wave modulation magnitude
(Fig. \ref{fig:analytical_solutions_ak0.1}a).
At the wave crest, (\ref{eq:k_short_exact}) yields the wavenumber modulation
that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of
\citet{longuet1960changes} for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
Although not immediately obvious from the expressions, the modulated wavenumber
(\ref{eq:k_short_lhs}) is conserved across the long-wave phase, whereas
the higher-order solution (\ref{eq:k_short_exact}) is not.
This is because combining the orbital velocities evaluated at $z=\eta_L$ and the
linearised conservation of wave crests in (\ref{eq:wavenumber}) creates excess
short-wave wavenumber.
The solution of (\ref{eq:wavenumber}) must be conservative across the long-wave
phase, as we will see later from the numerical simulations.

@@

\begin{equation}
\label{eq:gravity_modulation_linear_no_curvature}
\frac{\widetilde{g}}{g} = 1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}
\end{equation}
If we evaluate the orbital accelerations at the mean water level $z=0$,
this further simplifies to:

\begin{equation}
\label{eq:gravity_modulation_linear_no_curvature_mean_level}
\frac{\widetilde{g}}{g} = 1 - \varepsilon_L \cos{\psi}
\end{equation}
which is the form used by \citet{peureux2021unsteady}, for example.

@@

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{figures/fig_effective_gravities.png}
\caption{
  (a) Analytical solutions for the effective gravitational acceleration
  modulation by long waves, as function of long-wave phase, for $\varepsilon_L = 0.2$.
  Black is for linear wave evaluated at $z=0$,
  blue and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and without curvature effects,
  and orange and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and with curvature effects.
  (b) As in (a) but normalised by the modulation of the linear wave evaluated at $z=0$, $g_1$.
}
\label{fig:effective_gravities}
\end{figure}

It is instructive to also examine the mean, maximum, and minimum values of the
effective gravity as function of long-wave steepness $\varepsilon_L$
(Fig. \ref{fig:effective_gravities_mean_max_min}).
The long-wave averaged effective gravity of short waves remains equal to the
resting value of $g$ for all values of $\varepsilon_L$, except if the effects
of curvature are considered, in which case the average effective gravity is
reduced.
This reduction, in case of a long Stokes wave, is $\approx 3\%$ for
$\varepsilon_L = 0.2$ and $\approx 14\%$ for $\varepsilon_L = 0.4$.
As the effective gravity minima and maxima occur at crests and troughs,
respectively, the curvature affects the phase-averaged values but not the extremes.
As the slope effects on the effective gravity exist only away from crests and
troughs, they may be neglected in the steady solutions where modulation is
typically evaluated at the long-wave crests. 
The effective gravity reduction at the crests is $\approx 11\%$, $25\%$, and $66\%$
for $\varepsilon_L = 0.1$, $0.2$, and $0.4$, respectively.
These are similar to the fully-nonlinear numerical estimates of
\citet{longuet1986eulerian,longuet1987propagation}.
Note that these are strictly Eulerian values, evaluated at fixed positions in
space.
As \citet{longuet1986eulerian} pointed out, the effective gravity in the
Lagrangian frame (\textit{i.e.}, following the surface particles) is smoother
and less amplified because the moving particles spend a longer time riding near
the crests and less time returning through the troughs.
However, when we solve the short-wave crest and action conservation equations
in an Eulerian frame, it is the Eulerian effective gravity that must be applied
to conserve the short-wave properties across the long-wave phase.

@@

\begin{equation}
\label{eq:wave_amplitude_modulation}
\dfrac{\widetilde{a}}{a} = \sqrt{
  \dfrac{g}{\widetilde{g}}
  \dfrac{\widetilde{\sigma}}{\sigma}
  \dfrac{\widetilde{N}}{N}}
=
  \left( \dfrac{\widetilde{k}}{k} \right)^{\frac{1}{4}}
  \left( \dfrac{\widetilde{N}}{N} \right)^{\frac{1}{2}}
  \left( \dfrac{g}{\widetilde{g}} \right)^{\frac{1}{4}}
\end{equation}
To the first order in $\varepsilon_L$, and using (\ref{eq:k_short_exact})
and (\ref{eq:wave_action_modulation}), the above simplifies to the result of
\citet{longuet1960changes}:

\begin{equation}
\label{eq:wave_amplitude_modulation_order1}
\dfrac{\widetilde{a}}{a} = 
  \left( \dfrac{g}{\widetilde{g}} \right)^{\frac{1}{4}}
  \left( 1 + \varepsilon_L \right)^{\frac{3}{4}}
  \approx \frac{(1 + \varepsilon_L)^{\frac{3}{4}}}{(1 - \varepsilon_L)^{\frac{1}{4}}}
  \approx 1 + \varepsilon_L
\end{equation}

The amplitude and steepness modulation of the short waves as a function of the
long-wave phase for $\varepsilon_L = 0.1$ are compared to the
\citet{longuet1960changes} solutions in Fig.
\ref{fig:analytical_solutions_ak0.1} panels (c) and (d), respectively.
Also shown are the modulation of the short-wave intrinsic frequency and phase
speed in panels (e) and (f), respectively.
As the wavenumber and gravity modulations are out of phase, with the latter
being somewhat larger in magnitude, the overall intrinsic frequency of the
short waves is reduced across the entire long-wave phase, with highest reduction
at the long-wave crests.
For the short-wave phase speed modulation, the wavenumber and gravity modulations
work in tandem and cause the short-wave phase speed to be reduced on the long-wave
crests and increased in the troughs.
We discuss the implications of this effect later in \S\ref{section:discussion}
in the context of wind input to short waves.

Finally, the short-wave steepness modulation follows from (\ref{eq:k_short_exact})
and (\ref{eq:wave_amplitude_modulation}):

\begin{equation}
\label{eq:ak_modulation}
\dfrac{\widetilde{ak}}{ak} = 
  \left( \dfrac{\widetilde{k}}{k} \right)^{\frac{5}{4}}
  \left( \dfrac{\widetilde{N}}{N} \right)^{\frac{1}{2}}
  \left( \dfrac{g}{\widetilde{g}} \right)^{\frac{1}{4}}
\end{equation}
The short-wave steepness modulation at the long-wave crests is thus largely
controlled by the wavenumber modulation (5/8, or, 62.5\%), followed by the
convergence of wave action (1/4, or, 25\%), and by least amount by the
effective gravity reduction (1/8, or, 12.5\%) (Fig. \ref{fig:modulation_contributors}).
(\ref{eq:ak_modulation}) and Fig. \ref{fig:modulation_contributors} demonstrate
that more than half of the steepness modulation can be captured by the
wavenumber modulation alone.
Further, considering the wavenumber and wave action modulations, while
neglecting the gravity modulation, captures $\approx$88\% of the steepness
modulation at the long-wave crests.
This result may be useful for the interpretation of remote sensing products
that resolve some but not all of the short-wave modulation effects.
Finally, the relative contributions of different modulation factors remain
mostly constant as $\varepsilon_L$ increases, with only a mild increase of the
effective gravity modulation contribution, and mild increases of the
wavenumber and wave action modulation contributions.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/fig_modulation_contributors.png}
\caption{
  Contributions of short-wave wavenumber, action, and effective gravity
  modulations to the steepness modulation, as function of the long-wave
  steepness $\varepsilon_L$.
  Panel (a) shows each modulation factor (colour) and their product (black),
  and panel (b) shows their relative contributions in percentage, calculated as
  the ratio of the logarithm of each modulation factor to the logarithm of the
  product of all modulations.
}
\label{fig:modulation_contributors}
\end{figure}

@@

Why does this unsteady growth of short-wave action occur?
\citet{peureux2021unsteady} explain that the change in advection velocity due to
the modulation of the short-wave group speed yields an additional amplification
of the short-wave action.
That is, the inhomogeneity term $N \partial C_g / \partial x$ is
responsible for the instability.
However, the instability only occurs if the short waves are initialised as a
uniform field.
If they are initialised instead as a periodic function of the long-wave phase
such that their wavenumber (and optionally, action) is higher on the crests
and lower in the troughs, akin to the prior steady solutions, the instability
vanishes and the short-wave modulation pattern locks-in to the long-wave
crests.

@@

Why is it, then, that the solution for short-wave modulation is not only
quantitatively, but also qualitatively, different depending on the short-wave
initial conditions?
It may at first seem that there is something special about the uniform initial
conditions that makes its solutions grow unsteadily and indefinitely.
\citet{peureux2021unsteady} described this scenario as "...the sudden appearance
of a long wave perturbation in the middle of a homogeneous short wave field".
However, it is not obvious whether such a scenario is realistic as waves tend to
disperse into groups, and the time scales of incoming swell fronts are much
longer than those of short-wave generation by wind.
To test whether the sudden onset of long-wave forcing causes the unsteady growth,
we repeat the numerical simulation of Fig.
\ref{fig:modulation_3panel_infinite}, except for a linear ramp applied to $a_L$
(and consequently, the other long-wave properties including the orbital
velocities and the local steepness) for the first five periods:
$a_L(t) = min(a_L, a_L t / \left(5 T_L\right))$, where $T_L$ is the long-wave period.
This result is shown in Fig. \ref{fig:modulation_3panel_ramp}.
The key distinction that arises with the introduction of the ramp is that the
modulation pattern of short waves is now locked-in to the long-wave crests,
and their indefinite steepening vanishes.
The stabilisation effect of the long-wave ramp is more readily assessed in
Fig. \ref{fig:unsteady_growth_timeseries}, which shows the maximum short-wave
steepness modulation as function of time over 30 long-wave periods, for the
infinite long-wave train case and the case of applying a linear ramp.
Comparing the wave action propagation ($C_g \partial N / \partial x$) and
inhomogeneity ($N \partial C_g / \partial x$) tendencies reveals that
after only one long-wave period, the propagation and inhomogeneity tendencies
are unbalanced and amplify the steepening of short waves
(Fig. \ref{fig:inhomogeneity_tendencies}).
\citet{peureux2021unsteady} correctly pointed out that the unsteady growth of
short-wave steepness occurs due to the resonance of the inhomogeneity
tendency with the short-wave action perturbation.
However, in the case of a gentle ramp-up of the long-wave forcing, the
propagation and inhomogeneity tendencies mostly balance each other and allow
the short-wave steepening to lock-in to the long-wave crests.

@@

In this paper we revisited the problem of short-wave modulation by long waves
from a linear wave theory perspective.
Using the linearised wave crest and action conservation equations coupled with
nonlinear solutions for the effective gravity, we derived steady analytical
solutions for the modulation of short wave wavenumber, amplitude, steepness,
intrinsic frequency, and phase speed.
The latter two have not been previously examined in the context of short-wave
modulation, but may be important for correctly calculating the wind input into
short waves, which is proportional to the wind speed relative to the wave
celerity.
The approximate steady solutions yield higher wavenumber and amplitude
modulation than that of \citet{longuet1960changes}, and somewhat lower than 
the numerical solutions of \citet{longuet1987propagation} and
\citet{zhang1990evolution}.
This is mainly due to the analytical solution requiring the crest and action
conservation equations to be linearised, and the nonlinearity arising due
to the evaluation of short waves on the surface of the long wave (as opposed to
the mean water level) and due to nonlinear effective gravity.
In the steady solutions, the short-wave steepness modulation is dominated by
the wavenumber modulation (5/8), followed by the wave action modulation (1/4),
and is least affected by the effective gravity modulation (1/8).

The numerical solutions of the full wave crest and action conservation equations
are consistent with the analytical solutions for $\varepsilon_L \lesssim 0.2$,
and are similar in amplitude to the numerical solutions of
\citet{longuet1987propagation} and \citet{zhang1990evolution}.
The simulations also reveal that the unsteady growth of short-wave steepness
reported by \citet{peureux2021unsteady} is due to the sudden appearance of
long waves in a homogeneous short-wave field, a scenario that is unlikely to
occur in the open ocean.
A more likely condition such as that of long-wave groups, are only mildly
destabilising for the short waves.
These results suggest that for mildly-to-moderately steep short waves,
the analytical steady solutions presented here are sufficient to describe the
modulation of short waves by long waves.
It may be worthwhile to revisit a more systematic inclusion of the hydrodynamic
modulation effects in phase-averaged spectral wave models, beyond the simple
cumulative mean squared slope for wave dissipation, but also for modulating
the dispersion and wind input into short waves.

\backsection[Acknowledgements]{
I am thankful to Nathan Laxague, Fabrice Ardhuin, Nick Pizzo, and Peisen Tan
for the discussions that helped improve this paper.
}

\backsection[Funding]{
  Milan Curcic was partly supported by the National Science Foundation
  grant AGS-1745384, Transatlantic Research Partnership, and the Office of Naval
  Research Coastal Land Air-Sea Interaction DRI.
}
